The core transcriptome of mammalian placentas and the divergence of expression with placental shape

### 2016-12-05 manuscript/manuscript.Rnw
fix acronyms
switch from homologs to similar to
 - We aren't claiming that the Syncytins are separated by a speciation
   event (especially as they are endogenous retroviruses)

%\documentclass[english,12pt]{article}
%\documentclass[5p]{elsarticle_dla}
\documentclass{elsarticle_dla}
\usepackage{fontspec}
\setmainfont{FreeSerif}
\setsansfont{FreeSans}
\setmonofont{FreeMono}
\setmathrm{FreeSerif}
\setmathsf{FreeSans}
\setmathtt{FreeSerif}
%\usepackage[letterpaper,left=1in,right=1in,top=1in,bottom=1in,headheight=14pt]{geometry}
%\usepackage[letterpaper]{geometry}
%\usepackage{url}
\usepackage{graphicx}
%\usepackage{rotating}
%\usepackage{wrapfig}
\usepackage{booktabs}
%\usepackage{float}
\usepackage{multirow}
%\usepackage{setspace}
\usepackage[macros=false]{acro}
\usepackage{array}
\usepackage{longtable}
\usepackage{amsmath}
\usepackage[backend=biber,natbib=true,hyperref=true,citestyle=numeric-comp,style=nature,autocite=inline]{biblatex}
\addbibresource{references.bib}
\usepackage[hyperfigures,bookmarks,colorlinks,linkcolor=black,urlcolor=black,citecolor=blue,breaklinks=true]{hyperref}
% Floats at end for journals
% \usepackage[nolists,figuresfirst,nomarkers,heads]{endfloat}
% \renewcommand{\efloatseparator}{\mbox{}}
%\AtBeginFigures{\newpage \renewcommand{\efloatseparator}{\mbox{}}\section*{Figure Legends}}
%\AtBeginTables{\newpage \section*{Tables}\renewcommand{\efloatseparator}{\mbox{}}}
% \renewcommand{\includegraphics}[2][]{}
\usepackage[nomargin,inline,draft]{fixme}
\newcommand{\DLA}[1]{\textcolor{red}{\fxnote{DLA: #1}}}
\newcommand{\DEW}[1]{\textcolor{purple}{\fxnote{DEW: #1}}}
\newcommand{\PP}[1]{\textcolor{orange}{\fxnote{PP: #1}}}
% \usepackage[x11names,svgnames]{xcolor}
% \usepackage{texshade}
% \usepackage{tikz}
\usepackage{nameref}
\usepackage{xspace}
% this package will stretch the font slightly if necessary.
% \usepackage{microtype}
\usepackage{zref-xr,zref-user}
\zxrsetup{toltxlabel}
\zexternaldocument*{supplemental_material}
\usepackage[capitalize]{cleveref}
\def\UrlBreaks{\do\.\do\@\do\\\do\/\do\!\do\_\do\|\do\;\do\>\do\]%
\do\)\do\,\do\?\do\&\do\'\do+\do\=\do\#%
\do\/%
}%
\input{common_latex}
\tolerance=5000
\emergencystretch=0.1pt

@@

Another annexin which is highly expressed in all term placentas is
\ac{ANXA5}, which was originally identified as a an anti-coagulant
protein, but is also involved in syncytium formation, membrane repair,
and reduction of spontaneous
abortion~\cite{Ueki.ea2012:Lossofmaternalannexin}. Other notable
syncytiotrophoblast fusion proteins, Syncytin-1 (\ac{ERVW-1}) and -2
(\ac{ERVFRD-1}) were not present in the core placenta, as they (or
their homologs) are not highly expressed in \dog, \bonobo, or \sheep,
nor was the Syncytin-2 receptor, \ac{MFSD2A} (\ac{FPKM} $\ge 10$
only in \human, \mouse, and \dog).

@@

Supplemental \cref{tab:shape_genes} shows the 50 genes which are most
significantly associated with the shape of the placenta, as
well as the shape in which the difference is most significant. The
most significant genes which were associated with diffuse placenta
shape include \ac{EREG}, a ligand of EGFR, \ac{UNC93A}, a
homolog of \ac{UNC93B1} which is involved in the balancing between
Tlr7 and Tlr9 immune response in
mice~\cite{Brinkmann.ea2007:interactionbetweenERmembrane}, and
\ac{WNT5A}, which is a member of the Wnt ligand family whose
signaling pathway has been implicated in motility in human trophoblast
cells~\cite{Gupta.ea2016:CellSignalingPathwaysInvolved}.

@@

While Syncytin-1 and -2 (\ac{ERVW-1} and \ac{ERVFRD-1}, respectively) are implicated
in the formation of the syncytiotrophoblast in
\human~\cite{Rawn.ea2008:evolutionregulationandfunction} and homologs
of \ac{ERVFRD-1} and \ac{ERVW-1} are widely expressed in most of the
placentas analyzed here, they are not expressed in \bonobo, \sheep,
and \dog, even though these mammals form
syncytiotrophoblasts~\cite{Mess.ea2006:Evolutionarytransformationsoffetal}.
Because the formation of syncytiotrophoblast is the stem state of
Eutheria, Afrotheria, Euarchontoglires, and
Laurasiatheria~\cite{Mess.ea2006:Evolutionarytransformationsoffetal},
and syncytiotrophoblast must be continuously replenished
the gene(s) responsible for cytotrophoblast fusion to form
syncytiotrophoblast should be highly expressed in the core
placenta. \ac{ANXA5} is known to be expressed at multiple stages of
placentation, and is capable of repairing membrane defects and
promoting the fusion of
cells~\cite{Carmeille.ea2015:Annexin-A5promotesmembraneresealing} in
the presence of Ca$^{2+}$. While there may be other genes which
are responsible for the positioning of cytotrophoblast cells in close
proximity to start the fusion process, \ac{ANXA5} is likely to
be essential for the completion of fusion and the maintenance of
membrane repair during the massive flux of nutrients and waste in the
placenta \cite{Carmeille.ea2015:Annexin-A5promotesmembraneresealing,Ueki.ea2012:Lossofmaternalannexin}.

>>>>

%\documentclass[english,12pt]{article}
%\documentclass[5p]{elsarticle_dla}
\documentclass{elsarticle_dla}
\usepackage{fontspec}
\setmainfont{FreeSerif}
\setsansfont{FreeSans}
\setmonofont{FreeMono}
\setmathrm{FreeSerif}
\setmathsf{FreeSans}
\setmathtt{FreeSerif}
%\usepackage[letterpaper,left=1in,right=1in,top=1in,bottom=1in,headheight=14pt]{geometry}
%\usepackage[letterpaper]{geometry}
%\usepackage{url}
\usepackage{graphicx}
%\usepackage{rotating}
%\usepackage{wrapfig}
\usepackage{booktabs}
%\usepackage{float}
\usepackage{multirow}
%\usepackage{setspace}
\usepackage[macros=false,single=true,single-form=long]{acro}
\usepackage{array}
\usepackage{longtable}
\usepackage{amsmath}
\usepackage[backend=biber,natbib=true,hyperref=true,citestyle=numeric-comp,style=nature,autocite=inline]{biblatex}
\addbibresource{references.bib}
\usepackage[hyperfigures,bookmarks,colorlinks,linkcolor=black,urlcolor=black,citecolor=blue,breaklinks=true]{hyperref}
% Floats at end for journals
% \usepackage[nolists,figuresfirst,nomarkers,heads]{endfloat}
% \renewcommand{\efloatseparator}{\mbox{}}
%\AtBeginFigures{\newpage \renewcommand{\efloatseparator}{\mbox{}}\section*{Figure Legends}}
%\AtBeginTables{\newpage \section*{Tables}\renewcommand{\efloatseparator}{\mbox{}}}
% \renewcommand{\includegraphics}[2][]{}
\usepackage[nomargin,inline,draft]{fixme}
\newcommand{\DLA}[1]{\textcolor{red}{\fxnote{DLA: #1}}}
\newcommand{\DEW}[1]{\textcolor{purple}{\fxnote{DEW: #1}}}
\newcommand{\PP}[1]{\textcolor{orange}{\fxnote{PP: #1}}}
% \usepackage[x11names,svgnames]{xcolor}
% \usepackage{texshade}
% \usepackage{tikz}
\usepackage{nameref}
\usepackage{xspace}
% this package will stretch the font slightly if necessary.
% \usepackage{microtype}
\usepackage{zref-xr,zref-user}
\zxrsetup{toltxlabel}
\zexternaldocument*{supplemental_material}
\usepackage[capitalize]{cleveref}
\def\UrlBreaks{\do\.\do\@\do\\\do\/\do\!\do\_\do\|\do\;\do\>\do\]%
\do\)\do\,\do\?\do\&\do\'\do+\do\=\do\#%
\do\/%
}%
\input{common_latex}
\tolerance=5000
\emergencystretch=0.1pt

@@

Another annexin which is highly expressed in all term placentas is
\ac{ANXA5}, which was originally identified as a an anti-coagulant
protein, but is also involved in syncytium formation, membrane repair,
and reduction of spontaneous
abortion~\cite{Ueki.ea2012:Lossofmaternalannexin}. Other notable
syncytiotrophoblast fusion proteins, Syncytin-1 (\ac{ERVW-1}) and -2
(\ac{ERVFRD-1}) were not present in the core placenta, as they (or
genes similar to them) are not highly expressed in \dog, \bonobo, or \sheep,
nor was the Syncytin-2 receptor, \ac{MFSD2A} (\ac{FPKM} $\ge 10$
only in \human, \mouse, and \dog).

@@

Supplemental \cref{tab:shape_genes} shows the 50 genes which are most
significantly associated with the shape of the placenta, as
well as the shape in which the difference is most significant. The
most significant genes which were associated with diffuse placenta
shape include \ac{EREG}, a ligand of EGFR, \ac{UNC93A}, a
gene similar to \ac{UNC93B1} which is involved in the balancing between
Tlr7 and Tlr9 immune response in
mice~\cite{Brinkmann.ea2007:interactionbetweenERmembrane}, and
\ac{WNT5A}, which is a member of the Wnt ligand family whose
signaling pathway has been implicated in motility in human trophoblast
cells~\cite{Gupta.ea2016:CellSignalingPathwaysInvolved}.

@@

While Syncytin-1 and -2 (\ac{ERVW-1} and \ac{ERVFRD-1}, respectively) are implicated
in the formation of the syncytiotrophoblast in
\human~\cite{Rawn.ea2008:evolutionregulationandfunction} and genes which are similar to 
\ac{ERVFRD-1} and \ac{ERVW-1} are widely expressed in most of the
placentas analyzed here, they are not expressed in \bonobo, \sheep,
and \dog, even though these mammals form
syncytiotrophoblasts~\cite{Mess.ea2006:Evolutionarytransformationsoffetal}.
Because the formation of syncytiotrophoblast is the stem state of
Eutheria, Afrotheria, Euarchontoglires, and
Laurasiatheria~\cite{Mess.ea2006:Evolutionarytransformationsoffetal},
and syncytiotrophoblast must be continuously replenished
the gene(s) responsible for cytotrophoblast fusion to form
syncytiotrophoblast should be highly expressed in the core
placenta. \ac{ANXA5} is known to be expressed at multiple stages of
placentation, and is capable of repairing membrane defects and
promoting the fusion of
cells~\cite{Carmeille.ea2015:Annexin-A5promotesmembraneresealing} in
the presence of Ca$^{2+}$. While there may be other genes which
are responsible for the positioning of cytotrophoblast cells in close
proximity to start the fusion process, \ac{ANXA5} is likely to
be essential for the completion of fusion and the maintenance of
membrane repair during the massive flux of nutrients and waste in the
placenta \cite{Carmeille.ea2015:Annexin-A5promotesmembraneresealing,Ueki.ea2012:Lossofmaternalannexin}.
